Generating Empirically Optimized Numerical Software from MATLAB Prototypes

### 2008-09-19 doc/papers/cgo08/paper.tex
done for tonight
git-svn-id: https://svn.mcs.anl.gov/repos/performance/orio@350 e7ad4b5f-b827-0410-872f-f7f4bc3d1efb

\usepackage{url}
\usepackage{html}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cite}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}

@@

$x = w + y + z$

BN: working on this

@@

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.36\textwidth}
\footnotesize
\begin{verbatim}
ATAX
in
  A : row matrix, x : vector
out
  y : vector
{
  y = A' * (A * x)
}

@@

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.36\textwidth}
\footnotesize
\begin{verbatim}
GEMVER
in
  A : column matrix,
  u1 : vector, u2 : vector,
  v1 : vector, v2 : vector,
  a : scalar, b : scalar,
  y : vector, z : vector
out
  B : column matrix, 
  x : vector, w : vector
{
  B = A + u1 * v1' + u2 * v2'
  x = b * (B' * y) + z
  w = a * (B * x)
}
\end{verbatim}
\end{minipage}
&
\begin{minipage}[b]{.6\textwidth}
\includegraphics[width=\textwidth]{figures/gemver.png}
\end{minipage}\\
\end{tabular}
\caption{Annotated MATLAB code (left) and wall-clock time in seconds (right) for the GEMVER operation using different code versions.}
\label{fig:gesummv}
\end{figure}

@@

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.36\textwidth}
\footnotesize
\begin{verbatim}
GESUMMV
in
  A : row matrix,
  B : row matrix,
  x : vector,
  a : scalar,
  b : scalar
out
  y : vector
{
  y = a * (A * x) + b * (B * x)
}

@@

\begin{figure}[htp]
\centering

\includegraphics[width=.8\textwidth]{figures/bicgkernel.png}
\caption{Wall-clock time (in seconds) for the BiCG Kernel operation using different code versions.}
\label{fig:bicgkernel}
\end{figure}

>>>>

\setlength{\textwidth}{6.45in}
\setlength{\oddsidemargin}{0.05in}
\setlength{\evensidemargin}{0.05in}

\setlength{\topmargin}{-0.5in}
\setlength{\textheight}{9in}

\usepackage{url}
\usepackage{html}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cite}
\usepackage{amsmath}
%\usepackage[margin=1in]{geometry}

@@

We optimized the performance for the VADD operation, $x = w + y + z$, on the Blue Gene/P at Argonne National Laboratory using the IBM XLC compiler, version 9.0. We measured the performance for two scenarios: using one core per processor and using all cores. In the single-core case, the codes were compiled with the following options: \texttt{-O3 -qstrict -qarch=450d -qtune=450 -qhot -qsmp=noauto}. For the multicore case, the only difference in the compilation was that for the non-Orio versions, the \texttt{-qsmp=auto} option was used.

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.25\textwidth}
\footnotesize
\begin{verbatim}
VADD
in
  w : vector,
  y : vector,
  z : vector
out
  x : vector
{
  x = w + y + z
}

\end{verbatim}
\end{minipage}
&
\begin{minipage}[b]{.6\textwidth}
\includegraphics[width=\textwidth]{figures/vadd_bgp.png}
\end{minipage}\\
\end{tabular}
\caption{Annotated MATLAB code (left) and MFLOP/s (right) for the VADD operation using different code versions.}
\label{fig:atax}
\end{figure}

The "Base" label designates the simple loop C implementation generated by the Matlab compiler, which was simply compiled with the options mentioned earlier and executed. We tested the performance of the IBM ESSL library and Goto BLAS for the two calls to \texttt{daxpy} used in the BLAS version of the code. Finally, we annotated the simple C loop version with memory alignment and loop unrolling directives, and then used Orio to select the best tuned version for the single-core and multicore scenarios. We observe that even for a very simple operation, such as the vector addition above, the compiler alone is unable to obtain the same level of performance as the empirically tuned versions. Furthermore, as expected, the BLAS performance was worse than that of the single loop implementation, however, selecting the correct compiler options was crucial for achieving good performance of the base version.

@@

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.3\textwidth}
\footnotesize
\begin{verbatim}
ATAX
in
  A : row matrix, x : vector
out
  y : vector
{
  y = A' * (A * x)
}

@@

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.3\textwidth}
\footnotesize
\begin{verbatim}
GEMVER
in
  A : column matrix,
  u1 : vector, u2 : vector,
  v1 : vector, v2 : vector,
  a : scalar, b : scalar,
  y : vector, z : vector
out
  B : column matrix,
  x : vector, w : vector
{
  B = A + u1 * v1' + u2 * v2'
  x = b * (B' * y) + z
  w = a * (B * x)
}
\end{verbatim}
\end{minipage}
&
\begin{minipage}[b]{.6\textwidth}
\includegraphics[width=\textwidth]{figures/gemver.png}
\end{minipage}\\
\end{tabular}
\caption{Annotated MATLAB code (left) and wall-clock time in seconds (right) for the GEMVER operation using different code versions.}
\label{fig:gesummv}
\end{figure}

@@

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.3\textwidth}
\footnotesize
\begin{verbatim}
GESUMMV
in
  A : row matrix,
  B : row matrix,
  x : vector,
  a : scalar,
  b : scalar
out
  y : vector
{
  y = a * (A * x) + b * (B * x)
}

@@

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.3\textwidth}
\footnotesize
\begin{verbatim}
BICG
in
  A : column matrix,
  p : vector, r : vector
out
  q : vector, s : vector
{
  q = A * p
  s = A' * r
}

\end{verbatim}
\end{minipage}
&
\begin{minipage}[b]{.6\textwidth}
\includegraphics[width=\textwidth]{figures/bicgkernel.png}
\end{minipage}\\
\end{tabular}
\caption{Annotated MATLAB code (left) and wall-clock time in seconds (right) for the BiCG kernel operation using different code versions.}
\label{fig:gesummv}
\end{figure}
